\documentclass[a4wide, 11pt]{article}
\usepackage{a4, fullpage, amssymb}
\setlength{\parskip}{0.3cm}
\setlength{\parindent}{0cm}
\newcommand\xor{\mathbin{\char `\^}}
% This is the preamble section where you can include extra packages etc.

\begin{document}

\title{MAlice Language Specification}

\author{Robert Kruszewski \and Piotr Bar \and Łukasz Koprowski}


\date{\today}         % inserts today's date

\maketitle            % generates the title from the data above

\section{Introduction}

This document constitutes the specification of MAlice programming language.
First the BNF of the language is given. The BNF is followed by explanation
of language constructs and their meaning, i.e. what different operators mean
and what is an expected behaviour of these. Further some internal workings of compiler
are explained as they would faciliate understanding particular decisions when
it comes to language design.

\section{BNF Grammar} 

\begin{verbatim}
      assignment  := id function_name type assign_term
                  | id function_name expression 

      assign_term := separator
                  | separator assignment

      expression  := expression | or_expression

      or_expression := xor_expression
                    | or_expression '|' xor_expression
      xor_expression := and_expression
                     | xor_expression '^' and_expression 
      and_expression := add_expression
                     | and_expression '&' add_expression
      add_expression := mul_expression
                     | add_expression '+' mul_expression
                     | add_expression '-' mul_expression
      mul_expression := unary_expression
                     | mul_expression '/' unary_expression
                     | mul_expression '%' unary_expression
      unary_expression := primitive_expression
                       | un_op primitive_expression

      primitive_expression := id | numeral | letter | '(' expression ')' | \epsilon

      type        := 'letter' | 'number'

      separator   := '.' | ',' | 'then' | 'or' | 'and' | 'too' | 'but'

      un_op       := '~'
\end{verbatim}

\section{Semantics}

\subsection{Functions}

MAlice has a couple of predefined functions which perform certain operations on ids.

\textbf{was} - defines an id
\textit{cat was a number} $\to$ \emph{int cat;}

\textbf{became} - assigns value to an id
\textit{cat became 3} $\to$ \emph{cat = 3;}


\textbf{drank} - decrements value
\textit{cat ate} $\to cat++;$


\textbf{ate} - increments value
\textit{cat drank} $\to cat--;$
\subsection{Types}

From BNF you can easily see that MAlice supports two types: number and letter. Once defined variable cannot be redifined.

\begin{enumerate}
     \item
     number is a 8-bit positive integer number. Due to this format restriction and lack of special way of representing big numbers this is also the limit to calculations results. However, nothing stops the programmer from implementing additional features by himself.  

     \item
     letter is a character and it occupies 8-bit of memory. Character literals are enclosed in apostrophes. The ASCII encoding is supported.

\end{enumerate}

\subsection{Operators}

Operators have their usual meaning as used in other programming languages. All arithmetic operations are defined in the subset of $\mathbb{N+}$ which is representable as a 8-bit integer. In case of an overflow, MAlice returns biggest number it can represent.

\begin{itemize}

    \item
    \textit{+} - represents addition of two numbers, i.e. $ 2 + 3 = 5 $

    \item
    \textit{-} - represents subtraction of two numbers, i.e. $ 5 - 1 = 4 $. If the result of a substration is negative, MAlice returns 0.

    \item
    \textit{/} - is division of two numbers which always returns a \emph{number}, i.e. $ 6 / 3 = 2 $. Bear in mind that compiler will treat $ 23 / 0 $ as valid expression and returns 136.

    \item
    \textit{\%} - means a modulo operator, i.e. $ 4 \% 3 = 1 $

    \item
    \textit{\textasciicircum{}} - represents bitwise xor of two numbers, $ 0111011 \xor  0110010 = 0001001 $

    \item
    \textit{\&} - is used to represent bitwise and operation between two numbers,  $ 1101 \& 0110 = 0100 $

    \item
    \textit{$|$} - represents bitwise or of two numbers, i.e. $ 10110 | 01001 = 10111 $

    \item
    \textit{$\sim$} - is a unary operator which gives the bitwise inverse, $ \sim(010) = 101 $
    
\end{itemize}

When it comes to precedence of operators it is given by the following sequence

\begin{center} $\sim$, /, \%, +, -, \&, \textasciicircum{}, $|$ \end{center}

Operators precedence decreases from left to right. Therefore \emph{\textasciicircum{}} has highes precedence and \emph{$|$} lowest.
All of the operators are only valid for numbers as they meanings for letters would be unintuitive and best if would be left to the programmer to make the decision.

%Finally for this document, if you want to include a reference
%then you put it into a \texttt{thebibliography\{...\}}
%environment (see below in source file) and then 
%cite it like this \cite{lamport94}
%(you will need to run \texttt{latex} twice to get it to process the citation),
%or you can use BibTex but that is probably overkill for now.

\begin{thebibliography}{9}

\bibitem{lamport94}
  Leslie Lamport,
  \emph{\LaTeX: A Document Preparation System}.
  Addison Wesley, Massachusetts,
  2nd Edition,
  1994.

\end{thebibliography}


\end{document}
